---
title: "Simple AB Testing"
author: "Haozhe Chen"
date: "3/25/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(scipen=999, digits = 3)
```

## Introduction {.smaller}

The classic binary-choice version of AB testing has been used in marketing for many years.

Consider an example

```{r}

library(pwr)

test1 <- tibble(Product = c("A","B"), Views = c(3450, 3000), Conversions = c(225, 250))

test1

```

Are the conversion rates equal?

## Simple test {.smaller}

Ho: conversion rates are equal   
H1: Not Ho  
   
```{r}

prop.test(c(225, 250), c(3450, 3000))

```

We reject Ho.  
   
BUT What does this mean?  The test is limited to the visitors at the time of the test.  And, the outcome was a p-value.  It doesn't tell you how likely it is that B is similar to A or better than A. It doesn't tell you how much better B is than A.  

## Get (at least a little) realistic {.smaller}

If you're doing this kind of work, it probably isn't your "first rodeo."

Let's restate the question in Bayesian terms:

<div style="margin-left: 30px; font-size: small">

A is the proposition.  B is the evidence.  
P(A) is the prior probability of the proposition.  
P(B)Is the prior proabability of the evidence.  
P(B|A) is the likelihood. 
P(A|B) is the posterior.

Posterior = Likelihood * P(Prior proposition) / P(prior Evidence)  

</div>



## Uniform prior 1 {.smaller}

```{r}

library(bayesAB)

A_binom <- rbinom(3450, 1, 0.065)
B_binom <- rbinom(3000, 1, 0.083)


plotBeta(1,1)

```

## Uniform prior 2 {.smaller}

```{r}
AB1 <- bayesTest(A_binom, B_binom,
                 priors = c('alpha' = 1, 'beta' = 1),
                 distribution = 'bernoulli')

print(AB1)

```

## Uniform prior 3 {.smaller}

```{r fig.width= 3.3, fig.height=2.7}

plot(AB1)

```


## Beta Prior 1 {.smaller}

```{r}

plotBeta(100,200)

```

## Beta Prior 2

```{r}

AB2 <- bayesTest(A_binom, B_binom,
                 priors = c('alpha' = 100, 'beta' = 200),
                 distribution = 'bernoulli')
print(AB2)

```

## Beta Prior 3 {.smaller}

```{r fig.width= 3.3, fig.height=2.7}

plot(AB2)

```
