---
title: "Probability Distributions"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Compute probabilties 
```{r}
# X ~ N(0.05, (0.10)^2)
mu_x <-0.05
sigma_x <- 0.1

# Pr(X > 0.10)
1-pnorm(0.1,mu_x,sigma_x)

# Pr(X < -0.10)
pnorm(-0.1,mu_x,sigma_x)

# Pr(-0.05 < X < 0.15)
pnorm(0.15,mu_x,sigma_x)-pnorm(-0.05,mu_x,sigma_x)

```

## Quantiles
```{r}
# 1%, 5%, 95% and 99% quantile (the quantile of CDF)
qnorm(c(0.01,0.05,0.95,0.99),mu_x,sigma_x)
```

## Probability densities
```{r}
# Compute the probability desity of the vector x in under a normal PDF
x <- seq(-0.25, 0.35, length.out = 100)
d1 <- dnorm(x,0.05,0.10)
d2 <- dnorm(x,0.025,0.05)

# Plot the normal curves
plot(x,d1,type='l',col='blue',ylab='Normal curves',ylim=c(0,8))
lines(x,d2,col='red')
# Add a plot legend
legend("topleft", legend = c("density1", "density2"), 
       col = c("blue", "red"), lty = 1)

```

## Determine the value-at-risk(VaR) of returns
Assume that the simple monthly return R is normally distributed with a mean 0.04 and a variance (0.09)^2. The initial wealth to be invested over the month is $100,000.
```{r}
# R ~ N(0.04, (0.09)^2) 
mu_R <- 0.04
sigma_R <- 0.09

# Initial wealth W0 equals $100,000
W0 <- 100000

# The 1% value-at-risk
# The loss in investment value that may occur over the next month with a 1% probability
qnorm(0.01,mu_R,sigma_R)*W0

# The 5% value-at-risk
# The loss in investment value that may occur over the next month with a 5% probability
qnorm(0.05,mu_R,sigma_R)*W0

```

Assume that the log monthly return r is normally distributed with a mean 0.04 and a variance (0.09)^2. The initial wealth to be invested over the month is 100,000. Remember the transmutations: $r_t=log(R_t+1); R_t=exp(r_t)-1$
```{r}
# r ~ N(0.04, (0.09)^2) 
mu_r <- 0.04
sigma_r <- 0.09

# The 1% value-at-risk of simple return 
(exp(qnorm(0.01,mu_r,sigma_r))-1)*W0

# The 5% value-at-risk simple return 
(exp(qnorm(0.05,mu_r,sigma_r))-1)*W0
```


































